|

### 2021-08-11 prova_knit_finale.Rmd
correzioni

\begin{definition}
A function $P(x,dy): (\mathcal{X}, \mathcal{B}(\mathcal{X}))\rightarrow [0,1]$ such that 
\begin{itemize}
    \item for all $x\in\mathcal{X}, P(x,\cdot)$ is a probability measure on $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})$, and 
    \item for all subsets $A\in\mathcal{B}(\mathcal{Y}),$ the map $x\mapsto P(x,A)$ is measurable in $\mathcal{B}(\mathcal{X})),$
\end{itemize}
is defined a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})).$
\end{definition}
Notice that for any random variables $X_1$ and $X_0=x_0,$ 
$$
    \mathbb{P}_1(dx_{0:1})=\mathbb{P}_0(dx_{0})P_1(x_0,dx_1),
$$
where $P_1(x_0,dx_1)$ is a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{X}, \mathcal{B}(\mathcal{X})).$ This remark, provides some intuition about the link between probability kernels and conditional probabilities. As a matter of fact, it is possible to show that for given $X_{0:1}$ with joint probability distribution $\mathbb{P}_1(dx_{0:1}),$ 
\begin{equation}\label{prob_kern_condit_prob}
    \mathbb{P}_1(X_1\in dx_1\mid X_0=x_0)=P_1(x_0,dx_1).
\end{equation}

@@

\section{State-Space Models}\label{sec_state_space}
We now define \textit{State-Space models} as Markov processes that are partially observed. Let $X_t\in\X$ and $Y_t\in\Y$ and let $\{(X_t,Y_t)\}$ be a stochastic process. To help intuition, suppose that $\{Y_t\}$ is observable, i.e., coincides with a sequence of observations, while $\{X_t\}$ is latent. Consider the initial distribution $\P_0(dx_0)$ and the probability kernels
$P_t(x_{t-1},dx_t)$ and $F_t(x_t,d_t)$, where $t=1,2,\dots$.
\
\begin{definition}\label{state-space-def}
The process $\{(X_t,Y_t)\}$ is a State-Space model if the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{align*}
  \P_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T}) &= \P_0(dx_0)\prod\limits_{t=1}^T P_t(x_{t-1},dx_t)\prod\limits_{t=1}^T F_t(x_t,dy_t)\\
  &= [\P_0(dx_0)F_0(x_0,y_0)]\prod\limits_{t=1}^T[P_t(x_{t-1},dx_t)F_t(x_t,dy_t)] \\
  &= \P_T(dx_{0:T})\prod \limits_{t=1}^T F_t(x_t,dy_t).
\end{align*}
\end{definition}
From the second inequality, it is apparent that $\{(X_t,Y_t)\}$ is a Markov process with initial distribution $[\P_0(dx_0)F_0(x_0,y_0)]$ and transition kernel $P_t(x_{t-1},dx_t)F_t(x_t,dy_t).$ The third equality, instead, shows that, $\{X_t\}$ has a marginal distribution $$\P_T(dx_{0:T})=\P_0(dx_0)\prod_{t=1}^T P_t(x_{t-1},dx_t),$$ meaning that $\{X_t\}$ is marginally distributed as a Markov process with initial distribution $\P_T(dx_{0:T})$ and transition kernel $P_t(x_{t-1},dx_t).$

>>>>

\begin{definition}
A function $P(x,dy): (\mathcal{X}, \mathcal{B}(\mathcal{X}))\rightarrow [0,1]$ such that 
\begin{itemize}
    \item for all $x\in\mathcal{X}, P(x,\cdot)$ is a probability measure on $(\mathcal{Y}, \mathcal{B}(\mathcal{Y}))$, and 
    \item for all subsets $A\in\mathcal{B}(\mathcal{Y}),$ the map $x\mapsto P(x,A)$ is measurable in $\mathcal{B}(\mathcal{X}),$
\end{itemize}
is defined a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{Y}, \mathcal{B}(\mathcal{Y})).$
\end{definition}
Notice that for any random variables $X_1$ and $X_0=x_0,$ 
$$
    \mathbb{P}_1(dx_{0:1})=\mathbb{P}_0(dx_{0})P_1(x_0,dx_1),
$$
where $P_1(x_0,dx_1)$ is a probability kernel from $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ to $(\mathcal{X}, \mathcal{B}(\mathcal{X})).$ This remark, provides some intuition about the link between probability kernels and conditional probabilities. As a matter of fact, it is possible to show that for given $X_{0:1}$ with joint probability distribution $\mathbb{P}_1(dx_{0:1}),$ 
\begin{equation}\label{prob_kern_condit_prob}
    \mathbb{P}_1(X_1\in dx_1\mid X_0=x_0)=P_1(x_0,dx_1).
\end{equation}

@@

\section{State-Space Models}\label{sec_state_space}
We now define \textit{State-Space models} as Markov processes that are partially observed. Let $X_t\in\X$ and $Y_t\in\Y$ and let $\{(X_t,Y_t)\}$ be a stochastic process. To help intuition, suppose that $\{Y_t\}$ is observable, i.e., coincides with a sequence of observations, while $\{X_t\}$ is latent. Consider the initial distribution $\P_0(dx_0)$ and the probability kernels
$P_t(x_{t-1},dx_t)$ and $F_t(x_t,dy_t)$, where $t=1,2,\dots$.
\
\begin{definition}\label{state-space-def}
The process $\{(X_t,Y_t)\}$ is a State-Space model if the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{align*}
  \P_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T}) &= \P_0(dx_0)\prod\limits_{t=1}^T P_t(x_{t-1},dx_t)\prod\limits_{t=1}^T F_t(x_t,dy_t)\\
  &= [\P_0(dx_0)F_0(x_0,y_0)]\prod\limits_{t=1}^T[P_t(x_{t-1},dx_t)F_t(x_t,dy_t)] \\
  &= \P_T(dx_{0:T})\prod \limits_{t=1}^T F_t(x_t,dy_t).
\end{align*}
\end{definition}
From the second inequality, it is apparent that $\{(X_t,Y_t)\}$ is a Markov process with initial distribution $[\P_0(dx_0)F_0(x_0,y_0)]$ and transition kernel $P_t(x_{t-1},dx_t)F_t(x_t,dy_t).$ The third equality, instead, shows that, $\{X_t\}$ has a marginal distribution $$\P_T(dx_{0:T})=\P_0(dx_0)\prod_{t=1}^T P_t(x_{t-1},dx_t),$$ meaning that $\{X_t\}$ is marginally distributed as a Markov process with initial distribution $\P_T(dx_{0:T})$ and transition kernel $P_t(x_{t-1},dx_t).$
